A meta-analysis of fungicide effects on mungbean powdery mildew

### 2021-12-28 01_TrialsForInclusionInAnalysis.Rmd
Clean linebreaks, <br>, reformat code, etc.

pacman::p_load(tidyverse, kableExtra, leaflet, htmltools, lubridate, here)

@@

## Import data  
**Import data from trials which the raw data was available**  
```{r import_dataRAW}
PM_MB_raw <-
   rbind(
      read.csv(
         here("cache/2010 PMmung Hermitage means.csv"),
         stringsAsFactors = FALSE
      ),
      read.csv(here("cache/2011 PMmung Herm means.csv"), stringsAsFactors = FALSE),
      read.csv(
         here("cache/2011 PMmung Kingaroy means.csv"),
         stringsAsFactors = FALSE
      ),
      read.csv(here("cache/AM1305_Goolhi_means.csv"), stringsAsFactors = FALSE),
      read.csv(
         here("cache/AM1304-MarysMount_means.csv"),
         stringsAsFactors = FALSE
      ),
      read.csv(
         here("cache/AM1303-Premer-Disease_means.csv"),
         stringsAsFactors = FALSE
      ),
      read.csv(
         here("cache/BB1305_Millmerran_means.csv"),
         stringsAsFactors = FALSE
      ),
      read.csv(here("cache/Herm_16_means.csv"), stringsAsFactors = FALSE),
      read.csv(here("cache/King_16_means.csv"), stringsAsFactors = FALSE),
      read.csv(here("cache/Fogerty_17_mean.csv"), stringsAsFactors = FALSE),
      read.csv(here("cache/Hermitage_17_mean.csv"), stringsAsFactors = FALSE),
      read.csv(here("cache/Wellcamp_18_mean.csv"), stringsAsFactors = FALSE),
      read.csv(here("cache/Hermitage_19_mean.csv"), stringsAsFactors = FALSE)
   )
write.csv(
   PM_MB_raw,
   file = here("cache/PM_Mungbean_SummaryOfTrialsWithRawData"),
   row.names = FALSE
)
```

@@

<br>
<br>
<br>

@@

**Retain trials including: Yield**  
Three trials omitted due to not reporting yields  
```{r TrialWithoutYield}
PM_MB_dat %>%
   filter(is.na(grain_yield.t.ha.)) %>%
   distinct(trial_ref, year, location)

@@

**Retain trials including: Disease severity**  
One more was trial removed as it did not report disease severity.  
```{r TrialsWithotDisease}
PM_MB_dat %>%
   filter(is.na(PM_final_severity)) %>%
   distinct(trial_ref, year, location)

PM_MB_dat <- PM_MB_dat %>%
   filter(!is.na(PM_final_severity))
```

<br>
<br>
<br>

**Retain trials including: fungicide application dates**  
No trials from our subset need to be removed for not reporting fungicide application dates.  
```{r FungicideApplicationDates}
PM_MB_dat %>%
   group_by(trial_ref) %>%
   summarise(No_Record_Of_Fungicide_Application_Dates = all(is.na(fungicide_application_1)),
             .groups = 'drop') %>%
   filter(No_Record_Of_Fungicide_Application_Dates)
```

<br>
<br>
<br>

**Retain trials including: fungicide dose**  
No trials from our subset need to be removed for not reporting fungicide dose.  
```{r FungicideDoseTrials}
PM_MB_dat %>%
   group_by(trial_ref) %>%
   summarise(No_Record_Of_Fungicide_Dose = all(is.na(dose_ai.ha)),
             .groups = 'drop') %>%
   filter(No_Record_Of_Fungicide_Dose)
```

<br>
<br>
<br>

**Exclude fungicides tested in too few trials**  
The meta-analysis should be focused on fungicides with the same mode of action.
Fungicides from the best represented group will be retained.
```{r Fungicides}
PM_MB_dat %>%
   group_by(fungicide_ai, trial_ref) %>%
   summarise() %>%
   count(sort = TRUE) %>%
   rename(Trials = n) %>%
   ggplot(aes(x = reorder(fungicide_ai, Trials), y = Trials)) +
   xlab("Fungicide active ingredient") +
   ylab("N Trials") +
   geom_col() +
   ggtitle(label = "Number of trials in which the\nspecified fungicide was used") +
   coord_flip()
```

@@

<br>
<br>
<br>

**Retain only fungicides with DMI action**
```{r filterDMI_trials}
DMI_Trials <-
   PM_MB_dat %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   distinct(trial_ref) %>%
   pull()

@@

Now remove any non-DMI treatments or controls  
```{r KeepOnlyDMI}
PM_MB_dat <-
   PM_MB_dat %>%
   filter(
      fungicide_ai == "control" |
         fungicide_ai == "tebuconazole" |
         fungicide_ai == "propiconazole"
   )
```

@@

## Identify variance
We need variance accompanying the mean observations for each treatment to successfully undertake a meta-analysis.
Let's investigate which of the included trials included variance for both response variables of interest, yield and disease severity.

@@

Remove one trial in Emerald which did not report disease variance or yield variance.
```{r remove_nil_var}
# Remove trials without variance for both Disease and yield for data_set `PM_MB_dat
PM_MB_dat <- 
   PM_MB_dat %>%
   mutate(contains_var = case_when(
      is.na(yield_error) &
         is.na(disease_error) ~ NA_character_,
      TRUE ~ trial_ref)) %>%
   filter(trial_ref %in% contains_var) %>%
   dplyr::select(-contains_var) 
```

@@

One trial will be removed from the yield data set `PM_dat_Y` for not reporting yield variance.  
```{r TrialsWithVariance}
PM_MB_dat %>%
   mutate(No_Var = is.na(yield_error)) %>%
   filter(No_Var == TRUE) %>%
   distinct(trial_ref, year, location)

@@

Two more trials will be removed for our disease severity data set `PM_dat_D` for not reporting yield variance.
```{r TrialsWithVariance_D}
PM_MB_dat %>%
   mutate(No_Var = is.na(disease_error)) %>%
   filter(No_Var == TRUE) %>%
   distinct(trial_ref, year, location)

@@

### Yield variance standardisation
```{r StandardiseYieldVariance}
PM_dat_Y <-
   PM_dat_Y %>%
   mutate(
      vi =
         case_when(
            Y_error_type == "stdev" ~ yield_error ^ 2,
            Y_error_type == "lsd (P=0.05)" ~
               (replicates * ((yield_error / 1.96) ^ 2) / 2)
         ),
      id = row_number(),
      spray_management = fct_relevel(spray_management, sort)
   ) %>%
   dplyr::select(-c(yield_error,
                    Y_error_type))

@@

### Disease severity variance standardisation
```{r StandardiseSeverityVariance}
PM_dat_D <-
   PM_dat_D %>%
   mutate(
      vi =
         case_when(
            D_error_type == "stdev" ~ disease_error ^ 2,
            D_error_type == "lsd (P=0.05)" ~
               (replicates * ((disease_error / 1.96) ^ 2) / 2)
         ),
      id = row_number(),
      spray_management = fct_relevel(spray_management, sort)
   ) %>%
   dplyr::select(-c(disease_error,
                    D_error_type))

@@

#### Save the cleaned data

>>>>

pacman::p_load(tidyverse, kableExtra, leaflet, htmltools, lubridate, here)
```

@@

## Import data

**Import data from trials which the raw data was available**

```{r import_dataRAW}
PM_MB_raw <-
   rbind(
      read.csv(
         here("cache/2010 PMmung Hermitage means.csv"),
         stringsAsFactors = FALSE
      ),
      read.csv(here("cache/2011 PMmung Herm means.csv"),
               stringsAsFactors = FALSE),
      read.csv(
         here("cache/2011 PMmung Kingaroy means.csv"),
         stringsAsFactors = FALSE
      ),
      read.csv(here("cache/AM1305_Goolhi_means.csv"),
               stringsAsFactors = FALSE),
      read.csv(
         here("cache/AM1304-MarysMount_means.csv"),
         stringsAsFactors = FALSE
      ),
      read.csv(
         here("cache/AM1303-Premer-Disease_means.csv"),
         stringsAsFactors = FALSE
      ),
      read.csv(
         here("cache/BB1305_Millmerran_means.csv"),
         stringsAsFactors = FALSE
      ),
      read.csv(here("cache/Herm_16_means.csv"),
               stringsAsFactors = FALSE),
      read.csv(here("cache/King_16_means.csv"),
               stringsAsFactors = FALSE),
      read.csv(here("cache/Fogerty_17_mean.csv"),
               stringsAsFactors = FALSE),
      read.csv(here("cache/Hermitage_17_mean.csv"),
               stringsAsFactors = FALSE),
      read.csv(here("cache/Wellcamp_18_mean.csv"),
               stringsAsFactors = FALSE),
      read.csv(here("cache/Hermitage_19_mean.csv"),
               stringsAsFactors = FALSE)
   )
write.csv(
   PM_MB_raw,
   file = here("cache/PM_Mungbean_SummaryOfTrialsWithRawData"),
   row.names = FALSE
)
```

@@

@@

### Retain trials including: Yield

Three trials omitted due to not reporting yields

```{r TrialWithoutYield}
PM_MB_dat %>%
   filter(is.na(grain_yield.t.ha.)) %>%
   distinct(trial_ref, year, location)

@@

###Retain trials including: Disease severity

One more was trial removed as it did not report disease severity.

```{r TrialsWithotDisease}
PM_MB_dat %>%
   filter(is.na(PM_final_severity)) %>%
   distinct(trial_ref, year, location)

PM_MB_dat <- PM_MB_dat %>%
   filter(!is.na(PM_final_severity))
```
### Retain trials including: fungicide application dates

No trials from our subset need to be removed for not reporting fungicide application dates.

```{r FungicideApplicationDates}
PM_MB_dat %>%
   group_by(trial_ref) %>%
   summarise(No_Record_Of_Fungicide_Application_Dates = all(is.na(fungicide_application_1)),
             .groups = 'drop') %>%
   filter(No_Record_Of_Fungicide_Application_Dates)
```

### Retain trials including: fungicide dose

No trials from our subset need to be removed for not reporting fungicide dose.

```{r FungicideDoseTrials}
PM_MB_dat %>%
   group_by(trial_ref) %>%
   summarise(No_Record_Of_Fungicide_Dose = all(is.na(dose_ai.ha)),
             .groups = 'drop') %>%
   filter(No_Record_Of_Fungicide_Dose)
```

### Exclude fungicides tested in too few trials

The meta-analysis should be focused on fungicides with the same mode of action.
Fungicides from the best represented group will be retained.

```{r Fungicides}
PM_MB_dat %>%
   group_by(fungicide_ai, trial_ref) %>%
   summarise() %>%
   count(sort = TRUE) %>%
   rename(Trials = n) %>%
   ggplot(aes(x = reorder(fungicide_ai, Trials), y = Trials)) +
   xlab("Fungicide active ingredient") +
   ylab("N Trials") +
   geom_col() +
   ggtitle(label = "Number of trials in which the\nspecified fungicide was used") +
   coord_flip() +
   theme_classic()
```

@@

### Retain only fungicides with DMI action

```{r filterDMI_trials}
DMI_Trials <-
   PM_MB_dat %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   distinct(trial_ref) %>%
   pull()

@@

Now remove any non-DMI treatments or controls.

```{r KeepOnlyDMI}
PM_MB_dat <-
   PM_MB_dat %>%
   filter(
      fungicide_ai == "control" |
         fungicide_ai == "tebuconazole" |
         fungicide_ai == "propiconazole"
   )
```

@@

## Identify variance

We need variance accompanying the mean observations for each treatment to successfully undertake a meta-analysis.
Let's investigate which of the included trials included variance for both response variables of interest, yield and disease severity.

@@

Remove one trial in Emerald which did not report disease variance or yield variance.

```{r remove_nil_var}
# Remove trials without variance for both Disease and yield for data_set `PM_MB_dat
PM_MB_dat <- 
   PM_MB_dat %>%
   mutate(contains_var = case_when(
      is.na(yield_error) &
         is.na(disease_error) ~ NA_character_,
      TRUE ~ trial_ref)) %>%
   filter(trial_ref %in% contains_var) %>%
   dplyr::select(-contains_var) 
```

@@

One trial will be removed from the yield data set `PM_dat_Y` for not reporting yield variance.

```{r TrialsWithVariance}
PM_MB_dat %>%
   mutate(No_Var = is.na(yield_error)) %>%
   filter(No_Var == TRUE) %>%
   distinct(trial_ref, year, location)

@@

Two more trials will be removed for our disease severity data set `PM_dat_D` for not reporting yield variance.

```{r TrialsWithVariance_D}
PM_MB_dat %>%
   mutate(No_Var = is.na(disease_error)) %>%
   filter(No_Var == TRUE) %>%
   distinct(trial_ref, year, location)

@@

## Yield variance standardisation

```{r StandardiseYieldVariance}
PM_dat_Y <-
   PM_dat_Y %>%
   mutate(
      vi =
         case_when(
            Y_error_type == "stdev" ~ yield_error ^ 2,
            Y_error_type == "lsd (P=0.05)" ~
               (replicates * ((yield_error / 1.96) ^ 2) / 2)
         ),
      id = row_number(),
      spray_management = fct_relevel(spray_management, sort)
   ) %>%
   dplyr::select(-c(yield_error,
                    Y_error_type))

@@

## Disease severity variance standardisation

```{r StandardiseSeverityVariance}
PM_dat_D <-
   PM_dat_D %>%
   mutate(
      vi =
         case_when(
            D_error_type == "stdev" ~ disease_error ^ 2,
            D_error_type == "lsd (P=0.05)" ~
               (replicates * ((disease_error / 1.96) ^ 2) / 2)
         ),
      id = row_number(),
      spray_management = fct_relevel(spray_management, sort)
   ) %>%
   dplyr::select(-c(disease_error,
                    D_error_type))

@@

## Save the cleaned data
